HBCC - ACC - STG - DREAM - FEATURECOUNTS - GRCh38 - CQN

### 2020-04-02 cmc-rnaseq-de/HBCC_ACC-Nirmala_FEATURECOUNTS_GRCh38_CQN_Sparse.Rmd

# Github link
thisRepo <- getRepo(repository = "kelshmo/analysis", ref="branch", refName='add-dream')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('cmc-rnaseq-de/',thisFileName))
```
### Data download
#### Obtain effective count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
downloadFile <- function(id){
  fread(synGet(id)$path, data.table = F)
}
downloadFile_version <- function(id , version){
  fread(synGet(id, version = version)$path, data.table = F)
}

@@

adjust.covars = c(adjust.covars, vars.selected)
```

#### Variables included in the model 
Covariates were added as fixed effects iteratively in three phases if model improvement by BIC criteria was observed in the majority of genes. Clinical, ancestry and sample-specific technical variables were tested for model improvement first. Covariates related to batch effects were next tested for model improvement and finally, RNASeq alignment-specific covariates.

>>>>

# Github link
thisRepo <- getRepo(repository = "kelshmo/analysis", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('cmc-rnaseq-de/',thisFileName))
```
### Data download
#### Obtain effective count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
downloadFile <- function(id){
  fread(synGet(id)$path, data.table = F)
}
downloadFile_version <- function(id , version){
  fread(synGet(id, version = version)$path, data.table = F)
}

@@

adjust.covars = c(adjust.covars, vars.selected)
```
#### Variables included in the model 
Covariates were added as fixed effects iteratively in three phases if model improvement by BIC criteria was observed in the majority of genes. Clinical, ancestry and sample-specific technical variables were tested for model improvement first. Covariates related to batch effects were next tested for model improvement and finally, RNASeq alignment-specific covariates.

### 2020-04-02 cmc-rnaseq-de/HBCC_ACC-STG_FEATURECOUNTS_GRCh38_CQN_Sparse.Rmd

# Get RNASeq QCmetadata
METADATA_QC_ID = 'syn16816488' 
ALL_USED_IDs = METADATA_QC_ID
metadata = downloadFile_version(METADATA_QC_ID, version = 18) %>%
  dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 
                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
                                      'rnaSeq_dissection:Brain_Region',
                                      'rnaSeq_isolation:RNA_Isolation_Batch',
                                      'rnaSeq_report:Library_Batch', 
                                      'rnaSeq_report:Flowcell_Batch', 'rnaSeq_report:Exclude?',
                                      'rnaSeq_report:Mapped_Reads',
                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
                                      'rnaSeq_report:Expression_Profiling_Efficiency',
                                      'rnaSeq_report:rRNA_Rate',
                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned',
                                      'rnaSeq_report:Transcripts_Detected')) %>% 
  dplyr::filter(Cohort %in% c("NIMH-HBCC-4")) %>% 
  dplyr::filter(`rnaSeq_dissection:Brain_Region` %in% c("ACC","STG"))

@@

# Get genotype ids from synapse
GENOTYPE_ID = 'syn16816490'
ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
  dplyr::inner_join(ANCESTRY)
```

@@

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, "SampleID")) %>%
  dplyr::mutate(label = SampleID) %>% 
  dplyr::mutate(label = ifelse((label %in% outliers), label, NA)) %>% 
  tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_') %>% 
  dplyr::mutate(Institution = "NIMH-HBCC")

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Institution, shape=Tissue, size=ageOfDeath))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~.)
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p
# Plot abberent distribution of logcpm counts
tmp1 = voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  tidyr::gather(SampleID, logCPM, -ensembl_gene_id) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID'))
# p = ggplot(tmp1 %>%
#              dplyr::filter(SampleID %in% indToRemove),
#            aes(x = logCPM, color = SampleID)) + geom_density() 
# p = p + theme_bw() %+replace% theme(legend.position = 'none') + facet_grid(Dx~., scale = 'free')
# p

>>>>

# Get RNASeq QCmetadata
METADATA_QC_ID = 'syn16816488' 
ALL_USED_IDs = METADATA_QC_ID

metadata = downloadFile_version(METADATA_QC_ID, version = 18) %>%
  dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 
                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
                                      'rnaSeq_dissection:Brain_Region',
                                      'rnaSeq_isolation:RNA_Isolation_Batch',
                                      'rnaSeq_report:Library_Batch', 
                                      'rnaSeq_report:Flowcell_Batch', 'rnaSeq_report:Exclude?',
                                      'rnaSeq_report:Mapped_Reads',
                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
                                      'rnaSeq_report:Expression_Profiling_Efficiency',
                                      'rnaSeq_report:rRNA_Rate',
                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned',
                                      'rnaSeq_report:Transcripts_Detected')) %>% 
  dplyr::filter(Cohort %in% c("NIMH-HBCC-4")) %>% 
  dplyr::filter(`rnaSeq_dissection:Brain_Region` %in% c("ACC","STG"))

@@

# Get genotype ids from synapse
GENOTYPE_ID = 'syn16816490'
ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
  dplyr::inner_join(ANCESTRY) %>% 
  dplyr::filter(is.na(`SNP_report:Exclude?`)) #Remove excluded samples here because there are duplicate individuals
```

@@

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, "SampleID")) %>%
    tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_') %>%
  dplyr::mutate(label = SampleID) %>% 
  dplyr::mutate(label = ifelse((label %in% outliers), label, NA)) %>% 
  tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_') %>% 
  dplyr::mutate(Institution = "NIMH-HBCC")

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Institution, shape=Tissue, size=ageOfDeath))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~.)
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p

# Plot abberent distribution of logcpm counts
tmp1 = voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  tidyr::gather(SampleID, logCPM, -ensembl_gene_id) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID') %>%
              tidyr::separate(Dx.Tissue, c('Dx', 'Tissue'), sep = '_'))

p = ggplot(tmp1 %>%
             dplyr::filter(SampleID %in% outliers),
           aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'top') + facet_grid(Dx+.~Tissue, scale = 'free')
p
